---
title: "MEE 621 - Activity 5 Data Import"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---

<h3> This is the answers of the exercise by Junmar Sales II </h3>
<h3> Email: 92922@usc.edu.ph </h3>
<h3> Dated 17 October 2020. </h3>
<h3>  You can also see this on my github pages: <a href="https://92922.github.io/dist/MEE621_DataImport.html" >MEE 621 - Activity 5 on Data Import</a>  </h3>


This report on the answers of the exercises outline Activity 4 on Tibbles all done in R Studio.

Let's load the prerequisite libraries, we have assumes that packages has been installed.

```{r}
library("tidyverse")
```

<h1> 11.2.2 Exercises </h1>

<h3> 
1. What function would you use to read a file where fields were separated with
“|”?
</h3> 

Use ```read_delim``` for example:

```{r}
read_delim("a|b|c\n1|2|3", delim = "|")
```

<h3>
2. Apart from file , skip , and comment , what other arguments do read_csv() and
read_tsv() have in common?
</h3>

We can use the following:
```col_names```, ```col_types```, ```locale```, ```na```, ```quoted_na```, 
```quote```, ```trim_ws```, ```skip```, ```n_max```, ```guess_max```, ```progress```

We were able to get it form this code:

```{r}
intersect(names(formals(read_csv)), names(formals(read_tsv)))
```

Looking into help, we got: 
<p>
- ```col_names``` and ```col_types``` are used to specify the column names and how to parse the columns

- ```locale``` is important for determining things like the encoding and whether “.” or “,” is used as a decimal mark.

- ```na``` and ```quoted_na``` control which strings are treated as missing values when parsing vectors

- ```trim_ws``` trims whitespace before and after cells before parsing

- ```n_max``` sets how many rows to read

- ```guess_max``` sets how many rows to use when guessing the column type

- ```progress``` determines whether a progress bar is shown.

In fact, the two functions have the exact same arguments:

```{r}
identical(names(formals(read_csv)), names(formals(read_tsv)))
```
</p>

<h3>
3. What are the most important arguments to ```read_fwf()``` ?
</h3>

Short answer is ```widths```. Some other people choose is ```col_positions``` which tells the function where data columns begin and end.

<h3>
4. Sometimes strings in a CSV file contain commas. To prevent them from causing problems
they need to be surrounded by a quoting character, like " or ' . By default,
read_csv() assumes that the quoting character will be " . What argument to
read_csv() do you need to specify to read the following text into a data frame?

                  "x,y\n1,'a,b'"
</h3>

```{r}
"x,y\n1,'a,b'"
```
```{r}
read_delim("x,y\n1,'a,b'", delim = ",", quote = "'")
```
For read_delim(), we will will need to specify a delimiter, in this case ",", and a quote argument.


However, this question is out of date. ```read_csv()``` now supports a quote argument, so the following code works.

```{r}
src <- "x,y\n1,'a,b'"
read_csv(src, quote = "'")
```

<h3>
5. Identify what is wrong with each of the following inline CSV files. What happens when you
run the code?

       
         read_csv("a,b\n1,2,3\n4,5,6")
         read_csv("a,b,c\n1,2\n1,2,3,4")
         read_csv("a,b\n\"1")
         read_csv("a,b\n1,2\na,b")
         read_csv("a;b\n1;3")
      
</h3>


```{r}
read_csv("a,b\n1,2,3\n4,5,6")
```

It needs 3rd column header, skips 3rd argument on each line.

To correct it: 

```{r}
read_csv("a,b\n1,2\n3,4\n5,6")
```

```{r}
read_csv("a,b,c\n1,2\n1,2,3,4")
```

- There is a missing 3rd value on 2nd line so currently makes NA.

To correct it: 

```{r}
read_csv("a,b,c\n1,2, 1\n2,3,4")
```


```{r}
read_csv("a,b\n\"1")
```

The 2nd value is missing and 2nd quote mark is missing (though quotes are unnecessary).

To correct it: 
```{r}
read_csv("a,b\n\"1\",\"2\"")
```



```{r}
read_csv("a,b\n1,2\na,b")
```

IT has character and numeric types,

```{r}
read_csv("a;b\n1;3")
```   


There is a need to make read_csv2() because is separated by semicolons.

To correct it: 
```{r}
read_csv2("a;b\n1;3")
```

<h1> 11.3.5 Exercises </h1>

<h3>
1. What are the most important arguments to locale() ?
</h3>

It depends on the parse_* type, e.g. 

- for double, e.g. locale = locale(decimal_mark = ",")

- for number, e.g. locale = locale(grouping_mark = ".")

- for character, e.g. locale = locale(encoding = "Latin1")

- for dates, e.g. locale = locale(lang = "fr")

The locale object has arguments to set the following:

- date and time formats: date_names, date_format, and time_format

- time zone: tz

- numbers: decimal_mark, grouping_mark

- encoding: encoding

<h3>
2. What happens if you try and set decimal_mark and grouping_mark to the same
character? What happens to the default value of grouping_mark when you set
decimal_mark to “,”? What happens to the default value of decimal_mark when you
set the grouping_mark to “.”?
</h3>


If the decimal and grouping marks are set to the same character, locale throws an error:


```{r}
locale(decimal_mark = ",")
```

If the grouping mark is set to a period, then the decimal mark is set to a comma

```{r}
locale(grouping_mark = ".")
```
*** can’t set both to be same–if you change one, other automatically changes

```{r}
parse_number("$135.435,45", locale = locale(grouping_mark = ".", decimal_mark = ","))
```
```{r}
parse_number("$135.435,45", locale = locale(grouping_mark = "."))
```

<h3>
3. I didn’t discuss the date_format and time_format options to locale() . What do
they do? Construct an example that shows when they might be useful.
</h3>

They provide default date and time formats. The readr vignette discusses using these to parse dates: since dates can include languages specific weekday and month names, and different conventions for specifying AM/PM

```{r}
locale()
```


THe date_format and time_format in locale() will let you set the default date and time formats
```{r}
parse_date("31 january 2015", format = "%d %B %Y")
```



```{r}
parse_date("31 january 2015", locale = locale(date_format = "%d %B %Y"))
```


<h3>
4. If you live outside the US, create a new locale object that encapsulates the settings for
the types of file you read most commonly.
</h3>

Yes, I lived outside the US. So I need to read the help page for locale() using ?locale to learn about the different variables that can be set.

As an example, consider Australia. Most of the defaults values are valid, except that the date format is “(d)d/mm/yyyy”, meaning that January 2, 2006 is written as 02/01/2006.

However, default locale will parse that date as February 1, 2006.

```{r}
parse_date("10/17/2020")
```

To correctly parse Philippine dates, define a new locale object.

```{r}
ph_locale <- locale(date_format = "%d/%m/%Y")
```

Using parse_date() with the au_locale as its locale will correctly parse our example date.

```{r}
parse_date("17/10/2020", locale = ph_locale)
```

<h3>
5. What’s the difference between read_csv() and read_csv2() ?
</h3>

The delimiter. The function read_csv() uses a comma, while read_csv2() uses a semi-colon (;). Using a semi-colon is useful when commas are used as the decimal point (as in Europe).


<h3>
6. What are the most common encodings used in Europe? What are the most common
encodings used in Asia? Do some googling to find out.
</h3>


Europe tends to use “%d-%m-%Y”
Asia tends to use “%d.%m.%Y”

UTF-8 is standard now, and ASCII has been around forever.

For the European languages, there are separate encodings for Romance languages and Eastern European languages using Latin script, Cyrillic, Greek, Hebrew, Turkish: usually with separate ISO and Windows encoding standards. There is also Mac OS Roman.

For Asian languages Arabic and Vietnamese have ISO and Windows standards. The other major Asian scripts have their own:

Japanese: JIS X 0208, Shift JIS, ISO-2022-JP
Chinese: GB 2312, GBK, GB 18030
Korean: KS X 1001, EUC-KR, ISO-2022-KR
The list in the documentation for stringi::stri_enc_detect() is a good list of encodings since it supports the most common encodings.

Western European Latin script languages: ISO-8859-1, Windows-1250 (also CP-1250 for code-point)
Eastern European Latin script languages: ISO-8859-2, Windows-1252
Greek: ISO-8859-7
Turkish: ISO-8859-9, Windows-1254
Hebrew: ISO-8859-8, IBM424, Windows 1255
Russian: Windows 1251
Japanese: Shift JIS, ISO-2022-JP, EUC-JP
Korean: ISO-2022-KR, EUC-KR
Chinese: GB18030, ISO-2022-CN (Simplified), Big5 (Traditional)
Arabic: ISO-8859-6, IBM420, Windows 1256
For more information on character encodings see the following sources.

The Wikipedia page Character encoding, has a good list of encodings.
Unicode CLDR project
What is the most common encoding of each language (Stack Overflow)
“What Every Programmer Absolutely, Positively Needs To Know About Encodings And Character Sets To Work With Text”, http://kunststube.net/encoding/.
Programs that identify the encoding of text include:

readr::guess_encoding()
stringi::str_enc_detect()
iconv
chardet (Python)


<h3>
7. Generate the correct format string to parse each of the following dates and times:
```{r}
d1 <- "January 1, 2010"
d2 <- "2015-Mar-07"
d3 <- "06-Jun-2017"
d4 <- c("August 19 (2015)", "July 1 (2015)")
d5 <- "12/30/14" # Dec 30, 2014
t1 <- "1705"
t2 <- "11:15:10.12 PM"
```
</h3>

Solution:

```{r}
parse_date(d1, "%B %d, %Y")
```


```{r}
parse_date(d2, "%Y-%b-%d")
```


```{r}
parse_date(d3, "%d-%b-%Y")
```

```{r}
parse_date(d3, "%d%.%b-%Y") #could use this alternatively
```


```{r}
parse_date(d4, "%B %d (%Y)")
```


```{r}
parse_date(d5, "%m/%d/%y")
```

```{r}
parse_time(t1, "%H%M")
```

```{r}
parse_time(t2, "%I:%M:%OS %p")
```

